### 2020-11-08 latex/acrobot.tex
Edited oscillation section to be more clear
Added alpha_dot for rotations, as per results from simulation.

Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]

To show that the first-order approximation (\ref{eqn:acrobot-muhat-approx}) is
gaining energy, we define the Poincar\'{e} section 
\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
in Figure \ref{fig:acrobot-Q} and we see that \(Q(\mu_0)\) is strictly positive,
with an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).

@@

From this point on, we do the exact same procedure as with oscillations.
MATLAB symbolic operations evaluate the nominal system's dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\alpha}_n &=  what
    .
\end{align}
Once again, \(\dot{\alpha}_n > 0\) so for small enough \(I\) we must also have
\(\dot{\alpha} > 0\).
We take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) with initial
\(\mu_0\).
We expand the solution of this into the same form as
(\ref{eqn:acrobot-muhat-approx}) where once again we have
\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
The dynamics for \(\hat{\mu}_1\) are
\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
    \dot{\hat{\mu}}_1 =
    \sign{\mu_0}\frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
        - c_\alpha\mu_0^2
    \right)}{
    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
    }
    .
\end{equation}

>>>>

Likewise, (\ref{eqn:khalil-perturbation-firstorder}) tells us that 
\(\hat{\mu}_1\) is the solution to
\begin{equation}\label{eqn:acrobot-mu1-dot}
    \begin{cases}
        \dot{\hat{\mu}}_1 = a(\alpha,\mu_0)\hat{\mu}_1 + b(\alpha,\mu_0)
        , \\
        \hat{\mu}_1(0) = 0
        .
    \end{cases}
\end{equation}
where
\begin{align*}
    a(\alpha,\mu_0) &= \pdiff{g}{\hat{\mu}}(\alpha,\mu_0,0)
    , \\
    b(\alpha, \mu_0) &= \pdiff{g}{I}(\alpha, \mu_0, 0)
    .
\end{align*}
These are difficult to compute by hand, so we resort to MATLAB symbolic
computation to reveal that
\begin{align*}
    a(\alpha,\mu_0) &= 0
    , \\
    b(\alpha,\mu_0) &= K \hat{b}(\alpha,\mu_0)
    ,
\end{align*}
where we define
\begin{align*}
    K &:= \frac{\bar{q}_a \sqrt{30m^2g l^3}}{15}
    , \\
    \hat{b}(\alpha,\mu_0) &:= \frac{
        \mu_0 |\sin(\alpha)| \left(
        5 c_{\mu_0} \cos(\mu_0 c_\alpha) - 8 \cos(\mu_0c_\alpha)^2 + 3
    \right)
    }{
    \sin(\mu_0)\sqrt{\cos(\mu_0c_\alpha) - c_{\mu_0}}
    }
    .
\end{align*}
Hence, the solution to (\ref{eqn:acrobot-mu1-dot}) is given by
\[
    \hat{\mu}_1(\alpha,\mu_0) =
    K \int \limits_0^\alpha \hat{b}(\sigma,\mu_0)d\sigma
    .
\]
Notice that \(\hat{b}(\alpha,\mu_0)\) is both even and 
\(\pi\)-periodic in \(\alpha\), so that the following properties hold:
\begin{equation}\label{eqn:bhat-integral-properties}
    \begin{split}
    \int \limits_0 ^ {N \pi} \hat{b}(\sigma,\mu_0) d\sigma
    &= N \int \limits_0^\pi \hat{b}(\sigma,\mu_0) d\sigma \, 
    \left(\forall N \in \mathbb{Z}_{> 0}\right)
    , \\
    \int \limits_{-\pi}^0 \hat{b}(\sigma,\mu_0) d\sigma
    &= \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    .
    \end{split}
\end{equation}
Suppose we start the system at \((\alpha,\mu) = (0,\mu_0)\). 
Then the new pseudo-radius after a half orbit where \(\alpha\) goes from \(0\)
to \(\pi\) (that is, when the orbit hits the negative \(q_u\)-axis)
is \(\rho = \hat{\mu}(\pi,\mu_0)\).
Hence, the change in pseudo-radius is \(\rho - \mu_0 = I\hat{\mu}_1(\pi,\mu_0)\).
Now look at the full orbit where \(\alpha\) goes from \(0\) to \(2\pi\). 
This lands on the positive \(q_u\)-axis with pseudo-radius
\(\tilde{\rho} = \mu_0 + I\hat{\mu}_1(2\pi,\mu_0)\).
As we know from (\ref{eqn:bhat-integral-properties}), we have
\(\hat{\mu}_1(2\pi,\mu_0) = 2 \hat{\mu}_1(\pi,\mu_0)\). 
This means the change in pseudo-radius after one entire oscillation is 
\(\tilde{\rho} - \mu_0 = 2I\hat{\mu}1(\pi,\mu_0) = 2 (\rho - \mu_0)\).
Hence, if we can show that \(\hat{\mu}_1(\pi,\cdot)\) is always positive, we
will guarantee that the pseudo-radius increases after every half oscillation.
This in turn guarantees \(q_u\) is increasing after every half oscillation, so
that eventually we will reach a rotation.

To show that the first-order approximation (\ref{eqn:acrobot-muhat-approx}) is
gaining energy, we define the Poincar\'{e} section 
\(P(\mu_0) := \hat{\mu}(\pi,\mu_0)\), which expands into
\[
    P(\mu_0) = \mu_0 + I \hat{\mu}_1(\pi,\mu_0) + R(\pi,I)
    .
\]
Let us define
\[
    Q(\mu_0) := \int \limits_0^\pi \hat{b}(\sigma,\mu_0)d\sigma
    ,
\]
so that \(\hat{\mu}_1(\pi,\mu_0) = K Q(\mu_0)\).
Note that \(K\) is a positive constant that takes into account the acrobot's
mass and length, while \(\hat{b}(\alpha,\mu)\) is adimensional;
that is, it is the same for every possible choice of acrobot.
Hence, \(Q(\mu_0)\) is the same for every acrobot.
If \(Q(\mu_0)\) is strictly positive, then \(\hat{\mu}_1(\pi,\mu_0)\) will be
positive no matter the values of \(m\), \(g\), \(l\), or \(\bar{q}_a\).

We numerically compute \(Q(\mu_0)\) for \(\mu_0 \in [10^{-5}, \pi - 10^{-5}]\)
in Figure \ref{fig:acrobot-Q}. 
We see that \(Q(\mu_0)\) is strictly positive and monotonically increasing, with
an asymptote at \(\mu_0 = \pi\). 
Simulations with smaller \(\mu_0\) vanish in MATLAB, so we believe that 
\(Q(\mu_0)\) is in fact positive for all \(\mu_0\).

@@

From this point on, we do the exact same procedure as with oscillations.
MATLAB symbolic operations evaluate the nominal system's dynamics as
\begin{align}\label{eqn:acrobot-rot-mu-dot-nom}
    \dot{\mu}_n &= 0
    , \\
    \label{eqn:acrobot-rot-alpha-dot-nom}
    \dot{\alpha}_n &=  \sign(\mu_0) 
    \frac{\sqrt{\mu^2 - 30m^2gl^3(1 - c_\alpha)}}{5ml^2}
    .
\end{align}
As is expected, \(\dot{\alpha}_n\) does not change sign because the acrobot is
rotating;
if \(\mu_0 > 0\), the orbit traces a curve going from \(\alpha = -\pi\) to 
\(\alpha = \pi\), and if \(\mu_0 < 0\) the curve goes in the reverse direction.
For small enough \(I > 0\) we must also have 
\(\dot{\alpha} = \sign{\mu_0}\) and can hence use \(\alpha\) as a new time
variable.

Take the time-scaled system \(\hat{\mu}(\alpha,I)\) with dynamics 
\(\dot{\hat{\mu}}(\alpha) = \frac{\dot{\mu}}{\dot{\alpha}}\) and initial
\(\mu_0\).
We expand the solution of this into the same form as
(\ref{eqn:acrobot-muhat-approx}) where once again we have
\(\hat{\mu}_0(\alpha) \equiv \mu_0\).
The dynamics for \(\hat{\mu}_1\) are
\begin{equation}\label{eqn:acrobot-rot-mu1-dot}
    \dot{\hat{\mu}}_1 =
    \frac{5m^2 g l^3 \bar{q}_a \left(
        m^2gl^3\left(18s_\alpha^2 + 30c_\alpha(1 - c_\alpha)\right)
        - c_\alpha\mu_0^2
    \right)}{
    |\mu_0|\sqrt{\mu_0^2 - 30m^2gl^3(1 - c_\alpha)}
    }
    .
\end{equation}
This is an even, \(2\pi\)-periodic function in \(\alpha\).
